### 2019-11-25 1-illusion/report/report.tex

\documentclass[journal]{IEEEtran}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{amsmath}
\usepackage{comment}
%\usepackage{cases}
%\usepackage{subeqnarray}
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{xcolor}
\usepackage{listings}
\lstset{
  basicstyle=\fontsize{9}{10}\selectfont\ttfamily,
  numbers=left,
  numberstyle= \tiny,
  keywordstyle= \color{ blue!70},
  commentstyle= \color{red!50!green!50!blue!50},
  frame=single,
  rulesepcolor= \color{ red!20!green!20!blue!20} ,
  escapeinside=``,
  xleftmargin=1.5em,xrightmargin=0em, aboveskip=1em,
  framexleftmargin=2em,
  showstringspaces=false,
  showtabs=false,
  breaklines=true
}
\lstdefinelanguage{Solidity}
{
  morekeywords={contract, mapping, address, uint, private, function, public, if, payable},
  morecomment=[l]{//},
  morestring=[b]"
}

\usepackage{multicol}
\usepackage{lipsum}
\usepackage{mathtools}
\usepackage{cuted}

\usepackage{amsmath}
\usepackage{extpfeil}
\usepackage{mathpartir}
\usepackage[mathscr]{eucal}

\usepackage{hyperref}
\usepackage{cleveref}

\crefformat{section}{\S#2#1#3} % see manual of cleveref, section 8.2.1
\crefformat{subsection}{\S#2#1#3}
\crefformat{subsubsection}{\S#2#1#3}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Motion Illusion Detection and Creation}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{
	Shangning~Xu,~\IEEEmembership{SJTU,}
	Jinyu~Li,~\IEEEmembership{SJTU,}
	Zhongye~Wang,~\IEEEmembership{SJTU,}
	Xiaoyi~Bao,~\IEEEmembership{SJTU,}
	and Chenxuan~Li,~\IEEEmembership{SJTU}
}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~13, No.~9, September~2014}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peer review papers.
\begin{IEEEkeywords}
Motion Illusion, Deep Neural Network, Generative Adversary Network
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:intro}
\IEEEPARstart{M}{otion} illusion is one type of optical illusion where people perceive motions from some static images.
Figure~\ref{fig:motion_example} shows some sample images with illusory motions.
\begin{figure}[h]
	\centering
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth]{fig/illusion_eg.jpg}
	\end{minipage}\hfill
	\begin{minipage}{0.48\linewidth}
		\includegraphics[width=\linewidth]{fig/illusion_eg.jpg}
	\end{minipage}
	\caption{Motion Illusion Examples}
	\label[]{fig:motion_example}
\end{figure}

Rest of the introduction \dots

\section{Motion Illusion Detection: Prednet \& Flownet}
\label{sec:detection}
In this section, we will address the detection problem of motion illusion in static images.
We first explore and re-implement illusion reproduction strategy devised by Watanable \textit{et al.} \cite{watanable2018illusory} in \cref{sec:detection_dnn}, where they use a combination of Prednet \cite{lotter2016deep}, a deep neural network, and optical flow analysis to identify illusory motions.
Based on such strategy, we extract several metrics to evaluate the intensity of illusory motion of images in \cref{sec:detection_measure}.

We embed entire detector into a neural network by implementing the optical flow analysis using the Flownet \cite{ilg2017flownet}.
Therefore, we could easily get back-propagated feedback on how to improve the illusory image.
The integrated detector with Flownet (IDF) serves as a static discriminator in the illusion generative model we propose in \cref{sec:generation}.

\subsection{Detecting Motion Illusion by DNN \& Optic Flow Analysis}
\label{sec:detection_dnn}
Watanable \textit{et al.} \cite{watanable2018illusory} manage to reproduce the illusory motion human perceive from images.
The key idea in their strategy is to simulate the visual perception of human.
Unlike machines, human do not treat the visual input to their eyes as separated images, but instead as a consecutive sequence of image, i.e., a video.
When human observing the image with illusory motions, they tend to believe the image is part of the dynamic video they see through eyes.
In the meantime, the illusion information embedded in images relies on such belief that the image is not static and can produce motions.
Therefore, human can perceive these information, while the computer treat images as numerous static pixels and would neglect those information.

Guided by this intuition, they first use Prednet, a DNN model that predicts future frames in a video given first few frames, to simulate the process of human thinking images as videos.
They then apply the optical flow analysis to the generated video to extract the optical flow vector at some critical pixels.
These vectors can be interpreted as velocity vectors of the illusory motion.
Figure~\ref{fig:scheme} shows the scheme diagram for such reproduction strategy.

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{fig/pred-flow-procedure.png}
	\caption{Prednet-Flow Scheme for Illusion Reproduction}
	\label{fig:scheme}
\end{figure}

% $$
% \left[\begin{matrix}
% 	x_1 & y_1 & v_{x1} & v_{y1} \\
% 	x_2 & y_2 & v_{x2} & v_{y2} \\
% 	\vdots & \vdots & \vdots & \vdots \\
% 	x_n & y_n & v_{xn} & v_{yn} \\
% \end{matrix}\right]
% $$

\subsection{Measuring Illusion Intensity}
\label{sec:detection_measure}

\subsection{Integrated Detector with Flownet}
\label{sec:detection_idf}

\section{Motion Illusion Creation: Asymmetric GAN}
\label{sec:generation}

\subsection{Motion Illusion Generator}
\label{sec:generation_generator}

\subsection{Sanity Constraint Criteria}
\label{sec:generation_criteria}

\section{Experiments}
\label{sec:experiments}

\subsection{Illusion Detection}
\label{sec:experiments_detection}

\subsection{Illusion Generation}
\label{sec:experiments_generation}

\section{Related Work}
\label{sec:related}

\section{Conclusion}
\label{sec:conclusion}
The conclusion of the paper \dots

\bibliographystyle{IEEEtran}
\bibliography{ref}

\end{document}